Grade Retention and Mixed Race

### 2022-09-07 products/paper/main.qmd
Make improvements to tables

tbl_summary %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race",
                  "n",
                  "clearly below expected grade", 
                  "mean age" )) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

@@

[^8]: It might seem odd to hold categorical variables at their mean, but because all categorical variables are technically entered into the models as 0/1 numeric dummy variables, their mean value can be calculated as the proportion of cases with a value of 1. This technique is frequently used in multilevel models to deal with grand and group mean centering of variables. <!-- TODO CITE raudenbush and bryk? -->

@@

tbl_resources %>%
  select(race, held_back_cf, fam_inc, own_home, m_ba, f_ba, f_born, par_fb,
         spk_eng_well, spk_eng_par) %>%
  mutate(held_back_cf=percent(held_back_cf/100, accuracy=0.1),
         own_home=percent(own_home/100, accuracy=0.1),
         m_ba=percent(m_ba/100, accuracy=0.1),
         f_ba=percent(f_ba/100, accuracy=0.1),
         f_born=percent(f_born/100, accuracy=0.1),
         par_fb=percent(par_fb/100, accuracy=0.1),
         spk_eng_well=percent(spk_eng_well/100, accuracy=0.1),
         spk_eng_par=percent(spk_eng_par/100, accuracy=0.1),
         fam_inc=dollar(fam_inc)) %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("race","behind expected grade, counterfactual","family income",
                  "own home","mother, 4-year college degree",
                  "father, 4-year college degree",
                  "foreign-born","either parent foreign-born",
                  "speak English well", 
                  "Both parents speak english well")) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3") %>%
  column_spec(2:10, width = "1.5cm") %>%
  landscape()
```

@@

|                   | Lower than both | Same as lower | Between | Same as higher | Higher than both |
|------------|:----------:|:----------:|:----------:|:----------:|:----------:|
| White/Asian       |        X        |               |         |                |                  |
| Black/Asian       |                 |   X (Asian)   |         |                |                  |
| Black/Latino      |                 |   X (Black)   |         |                |                  |
| White/Latino      |                 |   X (White)   |         |                |                  |
| Indigenous/Latino |                 |  X (Indig.)   |         |                |                  |
| Latino/Asian      |                 |   X (Asian)   |         |                |                  |
| White/Indigenous  |                 |               |    X    |                |                  |
| Indigenous/Asian  |                 |               |    X    |                |                  |
| Black/Indigenous  |                 |               |    X    |                |                  |
| Black/White       |                 |               |         |   X (Black)    |                  |

: Relative placement for biracial groups in comparison to monoracial constituent groups of counterfactual probability of being behind grade from resources alone.

>>>>

tbl_summary %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("Race",
                  "Sample size",
                  "Clearly behind expected grade", 
                  "Mean age" )) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

@@

[^8]: It might seem odd to hold categorical variables at their mean, but because all categorical variables are technically entered into the models as 0/1 numeric indicator variables, their mean value can be calculated as the proportion of cases with the indicated value. This technique is frequently used in multilevel models to deal with grand and group mean centering of variables. <!-- TODO CITE raudenbush and bryk? -->

@@

tbl_resources %>%
  select(race, held_back_cf, fam_inc, own_home, m_ba, f_ba, f_born, par_fb,
         spk_eng_well, spk_eng_par) %>%
  mutate(held_back_cf=percent(held_back_cf/100, accuracy=0.1),
         own_home=percent(own_home/100, accuracy=0.1),
         m_ba=percent(m_ba/100, accuracy=0.1),
         f_ba=percent(f_ba/100, accuracy=0.1),
         f_born=percent(f_born/100, accuracy=0.1),
         par_fb=percent(par_fb/100, accuracy=0.1),
         spk_eng_well=percent(spk_eng_well/100, accuracy=0.1),
         spk_eng_par=percent(spk_eng_par/100, accuracy=0.1),
         fam_inc=dollar(fam_inc)) %>%
  kbl(booktabs = TRUE,
      linesep = NULL, position = "p",
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("Race","Behind expected grade, counterfactual","Family income",
                  "Own home","Mother, 4-year college degree",
                  "Father, 4-year college degree",
                  "Foreign-born","Either parent foreign-born",
                  "Speak English well", 
                  "Both parents speak english well")) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  row_spec(index_biracial, background = "#D3D3D3") %>%
  column_spec(2:10, width = "1.5cm")
#%>%
#  landscape()
```

@@

```{r}
#| label: tbl-resource-summary
#| tbl-cap: Relative placement for biracial groups in comparison to monoracial constituent groups of counterfactual probability of being behind grade from resources alone.
race <- c("White/Asian","Black/Asian","Latino/Asian","White/Latino",
          "Black/Latino","Indigenous/Latino","White/Indigenous",
          "Black/Indigenous","Indigenous/Latino","Black/White")
lower_both <- c("X",rep("",9))
same_lower <- c("","X (Asian)","X (Asian)","X (White)","X (Black)","X (Indig.)",
                rep("",4))
between <- c(rep("",6),rep("X",3),"")
same_higher <- c(rep("",9),"X (Black)")
higher_both <- rep("",10)
tibble(race, lower_both, same_lower, between, same_higher, higher_both) %>%
 kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", 5)),
      col.names=c("Biacial","Lower than both","Same as lower","Between",
                  "Same as higher","Higher than both")) %>%
  kable_styling() %>%
  column_spec(1:6, width = "2cm")

```
